Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-12-08 paper_comparison/paper_comparison/paper_comparison.Rmd
update 2.1

It is especially common, particularly among new statistics students to report patterns when an experienced data analyst might quickly conclude that there are none. [ET: I don't know if the former statement is true? >> PL: change to students] Generally, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity. It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.

@@

(This section discusses the visual patterns data analysts expect to see and their implications.)

@@

mod <- poly_model(include_z = FALSE, sigma = 0.25)
dat <- mod$gen(300)
patchwork::wrap_plots(
  mod$plot(dat, theme = theme_light()) +
    ggtitle("A. Good residuals") + 
    xlab("fitted values") + ylab("residuals"),
  mod$plot(mod$set_prm("include_z", TRUE)$gen(300, 
    computed = select(dat, x, e)), 
    theme = theme_light()) + 
    ggtitle("B. Non-linearity") +
    xlab("fitted values") + ylab("residuals"),
  mod$plot(heter_model(b = 64)$gen(300, 
    computed = select(dat, x, e)), 
    theme = theme_light()) + 
    ggtitle("C. Heteroskedasticity") +
    xlab("fitted values") + ylab("residuals"),
  mod$plot(heter_model(b = 0, 
    e = rand_lognormal(sigma = 0.5))$gen(300, 
    computed = select(dat, x)), 
    theme = theme_light()) + 
    ggtitle("D. Non-normality") +
    xlab("fitted values") + ylab("residuals")
)
```

Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals. As shown in Figure \ref{fig:residual-plot-common-departures}, the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns. There are various types of departures from a good residual plot.

Non-linearity, heterskedasticity and non-normality are perhaps the three mostly checked departures.
Non-linearity is a type of model misspecification caused by failing to include higher order terms of the regressors in the regression equation. Any non-linear functional form of residuals on fitted values presented in the residual plot could be considered as an indicative of non-linearity. An example residual plot containing visual pattern of non-linearity is given at the top-right of Figure  \ref{fig:commondepartures}. One can clearly observe the "S-shape" from the residual plot as the cubic term is not captured by the misspecified model.

Heteroskedasticity refers to the presence of nonconstant error variance in a regression model. It is mostly due to the strict but false assumptions on the variance-covariance matrix of the error term. The usual pattern of heteroskedasticity on a residual plot is the inconsistent spread of the residuals across the horizontal axis. Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure \ref{fig:commondepartures}, or the "left-triangle" and "right-triangle" shape where the smallest variance occurs at one side of the horizontal axis.

@@

  table_dat %>%
  knitr::kable(col.names = c("Plot", "Departures", "Model structural", "Breusch-Pagan", "Shapiro–Wilk"),
               align = "llrrr",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Statistical significance testing for departures from good residuals for plots in Figure \\ref{fig:residual-plot-common-departures}. Shown are the $p$-values calculated for the conventional Model structural, the Breusch-Pagan and the Shapiro–Wilk tests. The good residual plot (A) is judged a good residual plot, as expected, by all tests. The non-linearity (B) is detected by all tests, as might be expected given the extreme structure.', 
               label = "example-residual-plot-table" )

>>>>

It is especially common, particularly among new data analytics students to report patterns when an experienced data analyst might quickly conclude that there are none. [ET: I don't know if the former statement is true? >> PL: change to students] Generally, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity. It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.

@@

@@

mod <- poly_model(include_z = FALSE, sigma = 0.25)
dat <- mod$gen(300)
patchwork::wrap_plots(
  mod$plot(dat, theme = theme_light()) +
    ggtitle("A. Good residuals") + 
    xlab("fitted values") + ylab("residuals"),
  mod$plot(mod$set_prm("include_z", TRUE)$gen(300, 
    computed = select(dat, x, e)), 
    theme = theme_light()) + 
    ggtitle("B. Non-linearity") +
    xlab("fitted values") + ylab("residuals"),
  mod$plot(heter_model(b = 64)$gen(300, 
    computed = select(dat, x, e)), 
    theme = theme_light()) + 
    ggtitle("C. Heteroskedasticity") +
    xlab("fitted values") + ylab("residuals"),
  mod$plot(heter_model(b = 0, 
    e = rand_lognormal(sigma = 0.5))$gen(300, 
    computed = select(dat, x)), 
    theme = theme_light()) + 
    ggtitle("D. Non-normality") +
    xlab("Fitted values") + ylab("Residuals")
)
```

Graphical summaries in which residuals are plotted against fitted values or other functions of the regressors that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals. As shown in Figure \ref{fig:residual-plot-common-departures}, the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns.

There are various types of departures from a good residual plot. Non-linearity, heterskedasticity and non-normality are perhaps the three mostly checked departures.

Non-linearity is a type of model misspecification caused by failing to include higher order terms of the regressors in the regression equation. Any non-linear functional form of residuals on fitted values presented in the residual plot could be considered as an indicative of non-linearity. An example residual plot containing visual pattern of non-linearity is given at plot B of Figure \ref{fig:residual-plot-common-departures}. One can clearly observe the "S-shape" from the residual plot as the cubic term is not captured by the misspecified model.

Heteroskedasticity refers to the presence of nonconstant error variance in a regression model. It is mostly due to the strict but false assumptions on the variance-covariance matrix of the error term. The usual pattern of heteroskedasticity on a residual plot is the inconsistent spread of the residuals across the horizontal axis. Visually, it sometimes results in the so-called "butterfly" shape as shown in the plot C of Figure \ref{fig:residual-plot-common-departures}, or the "left-triangle" and "right-triangle" shape where the smallest variance occurs at one side of the horizontal axis.

@@

  table_dat %>%
  knitr::kable(col.names = c("Plot", "Departures", "RESET", "Breusch-Pagan", "Shapiro–Wilk"),
               align = "llrrr",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Statistical significance testing for departures from good residuals for plots in Figure \\ref{fig:residual-plot-common-departures}. Shown are the $p$-values calculated for the conventional RESET, the Breusch-Pagan and the Shapiro–Wilk tests. The good residual plot (A) is judged a good residual plot, as expected, by all tests. The non-linearity (B) is detected by all tests, as might be expected given the extreme structure.', 
               label = "example-residual-plot-table" )

### 2022-12-08 paper_comparison/paper_comparison/paper_comparison.tex
update 2.1

It is especially common, particularly among new statistics students to
report patterns when an experienced data analyst might quickly conclude
that there are none. {[}ET: I don't know if the former statement is
true? \textgreater\textgreater{} PL: change to students{]} Generally,
one looks for noticeable departures from the model like non-linear
dependency or heteroskedasticity. It is also possible to conduct
hypothesis tests for non-linear dependence \citep{ramsey_tests_1969},
and use a Breusch-Pagan test \citep{breusch_simple_1979} for
heteroskedasticity.

@@

Graphical summaries in which residuals are plotted against fitted values
or other functions of the predictor variables that are approximately
orthogonal to residuals are referred to as standard residual plots in
\citet{cook1982residuals}. As shown in Figure
\ref{fig:residual-plot-common-departures}, the top-left panel is a good
residual plot with residuals evenly distributed at both sides of the
horizontal zero line showing no noticeable patterns. There are various
types of departures from a good residual plot.

Non-linearity, heterskedasticity and non-normality are perhaps the three
mostly checked departures. Non-linearity is a type of model
misspecification caused by failing to include higher order terms of the
regressors in the regression equation. Any non-linear functional form of
residuals on fitted values presented in the residual plot could be
considered as an indicative of non-linearity. An example residual plot
containing visual pattern of non-linearity is given at the top-right of
Figure \ref{fig:commondepartures}. One can clearly observe the
``S-shape'' from the residual plot as the cubic term is not captured by
the misspecified model.

Heteroskedasticity refers to the presence of nonconstant error variance
in a regression model. It is mostly due to the strict but false
assumptions on the variance-covariance matrix of the error term. The
usual pattern of heteroskedasticity on a residual plot is the
inconsistent spread of the residuals across the horizontal axis.
Visually, it sometimes results in the so-called ``butterfly'' shape as
shown in the bottom-left panel of Figure \ref{fig:commondepartures}, or
the ``left-triangle'' and ``right-triangle'' shape where the smallest
variance occurs at one side of the horizontal axis.

@@

\caption{\label{tab:example-residual-plot-table}Statistical significance testing for departures from good residuals for plots in Figure \ref{fig:residual-plot-common-departures}. Shown are the $p$-values calculated for the conventional Model structural, the Breusch-Pagan and the Shapiro–Wilk tests. The good residual plot (A) is judged a good residual plot, as expected, by all tests. The non-linearity (B) is detected by all tests, as might be expected given the extreme structure.}
\centering
\begin{tabular}[t]{llrrr}
\toprule
Plot & Departures & Model structural & Breusch-Pagan & Shapiro–Wilk\\
\midrule
A & None & 0.434 & 0.133 & 0.728\\
B & Non-linearity & \em{0.000} & \em{0.000} & \em{0.039}\\
C & Heteroskedasticity & 0.378 & \em{0.000} & \em{0.000}\\
D & Non-normality & 0.667 & 0.736 & \em{0.000}\\
\bottomrule
\end{tabular}
\end{table}

>>>>

It is especially common, particularly among new data analytics students
to report patterns when an experienced data analyst might quickly
conclude that there are none. {[}ET: I don't know if the former
statement is true? \textgreater\textgreater{} PL: change to students{]}
Generally, one looks for noticeable departures from the model like
non-linear dependency or heteroskedasticity. It is also possible to
conduct hypothesis tests for non-linear dependence
\citep{ramsey_tests_1969}, and use a Breusch-Pagan test
\citep{breusch_simple_1979} for heteroskedasticity.

@@

Graphical summaries in which residuals are plotted against fitted values
or other functions of the regressors that are approximately orthogonal
to residuals are referred to as standard residual plots in
\citet{cook1982residuals}. As shown in Figure
\ref{fig:residual-plot-common-departures}, the top-left panel is a good
residual plot with residuals evenly distributed at both sides of the
horizontal zero line showing no noticeable patterns.

There are various types of departures from a good residual plot.
Non-linearity, heterskedasticity and non-normality are perhaps the three
mostly checked departures.

Non-linearity is a type of model misspecification caused by failing to
include higher order terms of the regressors in the regression equation.
Any non-linear functional form of residuals on fitted values presented
in the residual plot could be considered as an indicative of
non-linearity. An example residual plot containing visual pattern of
non-linearity is given at plot B of Figure
\ref{fig:residual-plot-common-departures}. One can clearly observe the
``S-shape'' from the residual plot as the cubic term is not captured by
the misspecified model.

Heteroskedasticity refers to the presence of nonconstant error variance
in a regression model. It is mostly due to the strict but false
assumptions on the variance-covariance matrix of the error term. The
usual pattern of heteroskedasticity on a residual plot is the
inconsistent spread of the residuals across the horizontal axis.
Visually, it sometimes results in the so-called ``butterfly'' shape as
shown in the plot C of Figure \ref{fig:residual-plot-common-departures},
or the ``left-triangle'' and ``right-triangle'' shape where the smallest
variance occurs at one side of the horizontal axis.

@@

\caption{\label{tab:example-residual-plot-table}Statistical significance testing for departures from good residuals for plots in Figure \ref{fig:residual-plot-common-departures}. Shown are the $p$-values calculated for the conventional RESET, the Breusch-Pagan and the Shapiro–Wilk tests. The good residual plot (A) is judged a good residual plot, as expected, by all tests. The non-linearity (B) is detected by all tests, as might be expected given the extreme structure.}
\centering
\begin{tabular}[t]{llrrr}
\toprule
Plot & Departures & RESET & Breusch-Pagan & Shapiro–Wilk\\
\midrule
A & None & 0.434 & 0.133 & 0.728\\
B & Non-linearity & \em{0.000} & \em{0.000} & \em{0.039}\\
C & Heteroskedasticity & 0.378 & \em{0.000} & \em{0.000}\\
D & Non-normality & 0.667 & 0.736 & \em{0.000}\\
\bottomrule
\end{tabular}
\end{table}
